Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2016-10-07 sections0.07/method0.07.tex
adding K-means section

\section{METHOD}
\label{sec: method}
 \subsection{Self Organizing Maps}
 \label{sec: som}

 The self-organizing map is a clustering method which reduces the dimension of the data, usually to one or two dimensions (1D or 2D), while preserving topological features of the original dataset.
 The result of a SOM is a set of nodes (neurons) that are arranged in a 1D or 2D arrays \citep{Kohonen98}. 
 Each node may contain one or more samples from the input data.
 The distance between the nodes represents similarity or dissimilarity of the underlying samples, i.e., similar data are closer together in the array and the distance between two nodes is related to the dissimilarity of their samples.
 A weight vector ``\boldit{W}" with the same dimension as the input data is associated with each node and will be varied during the training process.
 This vector is the key factor in determining the position of the nodes in a map.
 \cite{Geach12} presented the application of the self-organized map and discussed its algorithm in detail.
 In this section we briefly discuss the algorithm of SOM, how we create our maps and a present a mock model which will help interpretation of our results.

@@

\subsection{Mock sample}

         \begin{figure}
            \begin{subfigure}[b]{0.5\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_dist.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_hits.png}
            \end{subfigure}
            \caption{SOM of the mock sample. In both plots, the axes show the position of the neurons. Hexagonal shapes represent the neurons. The top plot is a distance map. The grey cycle colours show the differences between the weights of each neuron with white being the minimum difference and black being the maximum one. The lower plot is a ``hits plot". It represents the number of samples in each neuron, where empty means zero hits. In this sample, the 27 galaxies are clustered in 4 groups, containing 8, 7, 6 and 6 galaxies respectively.}
            \label{fig: sample}
        \end{figure}

 To illustrate how self-organizing maps work, we create a mock sample of 27 galaxies.
 The mock sample contains two assigned attributes for each galaxy: the type (type 1 or type 2), and the star formation rate (high or low). %PB20160623: were you going to change type1/2 to A/B ?
 We generated a SOM of size $10 \times 10$ using the same initial values mentioned in Section ~\ref{sec: create_som}.

 Fig. ~\ref{fig: sample} shows the SOM of this mock sample. 
 The upper panel is the distance map. 
 The axes show the position of the neurons in the $10 \times 10$ network.
 The lower panel in Fig.~\ref{fig: sample} shows the hit map.
 On this map, similar to the distance map, the axes show the position of the neurons and the hexagonal shapes are the neurons.
 The number on the purple neurons shows the number of galaxies in that neuron.
 Colour coverage of neurons depends on the number of the hits on the sample.
 The neuron with the maximum number of hits is fully coloured while the empty neurons are left uncoloured (white).

Using this method, as expected, we are able to divide the mock sample galaxies into 4 distinct groups: Type 1 with high SFR, type 1 with low SFR, type 2 with high SFR and type 2 with low SFR. 
The upper panel in the Fig.~\ref{fig: sample} clearly shows this division.
In that plot, the upper half belongs to highly star-forming galaxies, while the lower half belongs to less star forming galaxies.
The left half of the plot is where type 1 galaxies belong, and type 2 galaxies reside in the right side.
Different shades of gray show the border between regions.
The lower panel in Fig.~\ref{fig: sample} shows that only 4 neurons (out of 100) are occupied. 
8 galaxies are type 1 galaxies with high SFR, 7 are type 2 galaxies with high SFR, 6 are type 1 galaxies with low SFR and the other 6 are type 2 galaxies with low SFR.

We can conclude that although each galaxy had the chance to occupy any of the neurons in the network, because of the similarity in the values of their attributes, they remained only in four groups.
This network is considered a training network and can be used to cluster any new dataset with similar entries.
Some neurons were left empty in this network because they never became a winner for data in the mock sample.
However, their weights have been updated in each iteration, and they could be a winner for data in the new dataset.

As we show in the following sections, with data from real galaxies there are more than two dimensions and two galaxies never have exactly the same information. 
If the network has enough neurons, the input data points would eventually separate from each other and cluster into smaller groups. 
However, if the input data has high similarity, the number of neurons must be much higher than the number of input samples to be able to separate the groups from each other. 
Therefore, it is up to the user to decide the similarity or dissimilarity between the input data based on the number of neurons.

>>>>

\section{METHOD}
\label{sec: method_somz}
 \subsection{Self Organizing Maps}
 \label{sec: som}

 The self-organizing map is a clustering method which reduces the dimensionality of the data, usually to one or two dimensions (1D or 2D), while preserving topological features of the original dataset.
 The result of an SOM is a set of nodes (neurons) that are arranged in a 1D or 2D arrays \citep{Kohonen98}. 
 Each node may contain one or more samples from the input data.
 The distance between the nodes represents similarity or dissimilarity of the underlying samples, i.e., similar data are closer together in the array and the distance between two nodes is related to the dissimilarity of their samples.
 A weight vector ``\boldit{W}" with the same dimension as the input data is associated with each node and will be varied during the training process.
 This vector is the key factor in determining the position of the nodes in a map.
 \cite{Geach12} presented the application of the self-organized map and discussed its algorithm in detail.
 In this section we briefly discuss the algorithm of SOM, how we create our maps.

@@
